Meta-Analysis of the `Ironic' Effects of Intergroup Contact

### 2021-06-07 manuscript/manuscript.tex
add starred references

\noindent Growing evidence suggests that intergroup contact, psychology's
most-researched paradigm for reducing prejudice, might have the `ironic'
effect of reducing support for social change in disadvantaged groups. We
conducted a preregistered meta-analysis of this effect across 97 studies
with 138 samples of 212,189 disadvantaged-group members. As predicted,
intergroup contact was, on average, associated with less perceived
injustice (\(r = -.07\)), collective action (\(r = -.06\)), and support
for reparative policies (\(r = -.07\)). Moderator analyses found that
associations were strongest in studies of adults that measured
intergroup contact directly and examined short-term migration or
post-colonial intergroup relations. After controlling for the positive
association of negative contact with support for social change, positive
contact was not associated with any outcome in studies that measured
both predictors. We close by highlighting open questions about the
relationship between intergroup contact and social change.

@@

It is ironic that psychology's most-researched paradigm for reducing
prejudice---intergroup contact---should hinder, rather than help, social
change. Hundreds of studies have confirmed that intergroup contact is
associated with less negative feelings toward outgroup members (for a
meta-analysis, see Pettigrew \& Tropp, 2006). Most studies have focused
on advantaged-group members' contact experiences with
disadvantaged-group members (Tropp \& Pettigrew, 2005), guided by the
idea that reducing advantaged-group members' prejudice will prevent
interpersonal discrimination. Reducing prejudice, however, is rarely
enough to overcome institutional and structural discrimination. Instead,
as the 2020 Black Lives Matter protests have shown, social change often
requires political mobilization by the disadvantaged. Critics have
argued that, by fostering harmony and reducing conflict, intergroup
contact might distract disadvantaged-group members from the injustice
they face and undermine their motivation for social change (Dixon et
al., 2005; Reicher, 2007; Wright, 2003). By reducing disadvantaged-group
members' opposition to injustice, intergroup contact might thus have the
`ironic' (Saguy et al., 2009), `sedative' (Çakal et al., 2011), or
`paradoxical' (Dixon et al., 2010) effect of hindering social change.

@@

Collective action---that is, any action directed at improving the
conditions of the disadvantaged ingroup (Wright et al., 1990)---is
crucial to achieving social change because advantaged-group members
rarely surrender their advantages without sustained pressure from
disadvantaged-group members (Blumer, 1958; for a review, see Dixon et
al., 2012). Researchers have argued that intergroup contact could
discourage disadvantaged-group members from engaging in collective
action by diminishing their perceived discrimination (Tropp et al.,
2012), by blurring boundaries between the disadvantaged ingroup and the
advantaged outgroup (Saguy et al., 2009), by reducing negative attitudes
toward the advantaged outgroup (Wright \& Lubensky, 2009), and by
quelling their anger (Hayward et al., 2018; Tausch et al., 2015). To the
extent that it discourages collective action, intergroup contact reduces
support for social change.

@@

Figure 1 shows a flow diagram illustrating our search strategy, study
selection, and data collection. Our preregistered search strategy
returned 2,610 unique records from electronic databases. Of these, we
excluded 2,379 (91\%) ineligible records after screening titles,
abstracts, and keywords. Of the remaining 231 records, we excluded 116
(50\%) ineligible records after reviewing full-text manuscripts. We
supplemented these records with 14 unpublished studies and 6 records
that cited at least three relevant works. Of 135 eligible studies, we
had to exclude 38 (28\%) studies for which we could not extract or
impute any relevant effect size. At each stage, we also excluded studies
that used the same data as another study. Our final sample comprised
effect sizes from \(J = 97\) studies spanning \(N = 212,189\)
participants in \(I = 138\) samples (for a complete list of studies, see
SOM-U).

Figure 2 provides a qualitative overview of the literature. Even though
most participants came from studies conducted in India (\(N = 104,639\))
and the United States (\(N = 60,983\); Figure 2a), most samples were
collected in North America (\(I = 48\)) and Europe (\(I = 41\)), with
few samples from South America and Africa (Figure 2b). Most samples
focused on relative inequalities that resulted from long-term migration
(\(I = 57\)), slavery (\(I = 19\)), colonization (\(I = 19\)), and
short-term migration (\(I = 14\)). With few exceptions, studies used
observational, cross-sectional survey designs. Samples sizes ranged from
64 to 49,764 (\(\textit{Mdn} = 283\)). Unlike most psychological
research, 2 in 5 samples used probability or representative sampling. Of
all samples, 54\% (\(I = 75\)) were collected with the intention to
examine the effects of intergroup contact on the outcomes considered in
this meta-analysis. Many studies instead examined acculturation
processes (Berry, 1997) and measured intergroup contact and perceived
injustice without hypothesizing that the former would influence the
latter. Together, these observations highlight both limitations and
strengths of the empirical literature.

@@

\textbf{\emph{Perceived Injustice.}} Across 202,741 participants from
124 samples in 85 studies, we found strong evidence for a weak
association (\(r = -.07, [-.10, -.04]\)) between intergroup contact and
perceived injustice, with \(>99.9\%\) of posterior samples for the mean
correlation coefficient falling below zero. We found evidence that
correlation coefficients varied across studies
(\(\tau_J = .14, [.12, .17]\)) and across samples within studies
(\(\tau_I = .08, [.05, .12]\)). Based on these analyses, we predicted
that 80\% of studies would result in correlation coefficients between
-.25 and .10 and that researchers would need sample sizes of at least
2,549 participants to find significant associations (\(\alpha = .05\),
two-sided) in 80\% of their studies.\footnote{Sample sizes are based on
  posterior predictions which implied that, for 80\% of studies, the
  absolute correlation coefficient would be \(|r| > .039\) for perceived
  injustice, \(|r| > .045\) for collective action, and \(|r| > .035\)
  for policy support.}

@@

\begin{figure*}
\centering
\caption{Estimated effect size as a function of cultural distance from the United States, with point estimates and uncertainty intervals for each country}
\includegraphics[scale=1]{../figures/figure-5}
\caption*{\textit{Note.} We had data from too few countries to reach firm conclusions about the direction of these associations for perceived injustice ($\beta = -0.34, [-1.41, 0.81]$; $\Pr (\beta < 0) = 73.5\%$), collective action ($\beta = -0.09, [-1.60, 1.37]$; $\Pr (\beta < 0) = 55.1\%$), and policy support ($\beta = -0.92, [-2.13, 0.55]$; $\Pr (\beta < 0) = 90.3\%$).}
\label{fig:f5}
\end{figure*}
\begin{figure*}
\centering
\caption{Results from the random-effects meta-regression tree analysis}
\includegraphics[scale=1]{../figures/figure-6}
\caption*{\textit{Note.} Posterior distributions for the estimated correlation coefficient in each leaf of the meta-regression tree, highlighting the proportion of posterior samples for which $r_\text{mean} < 0$.  S.-T. Migration = Short-Term Migration.}
\label{fig:f6}
\end{figure*}

@@

Second, we used meta-regression trees to discover interactions between
moderator variables that best explained heterogeneity in effect sizes
(Li et al., 2020). Figure 6 shows the resulting meta-regression that
explained more variance across samples than any individual moderator
(\(R^2 = 31\%\)). We found that intergroup contact was associated with
less perceived injustice only in studies that focused on adults and that
measured intergroup contact directly. Among these studies, this
association was stronger in settings in which the groups' relative
inequality stemmed from short-term migration or colonization
(\(r = -.18, [-.23, -.13]\)) than in other settings
(\(r = -.09, [-.12, -.05]\)).

@@

We ran three random-effects meta-analysis models estimating the partial
correlations of positive and negative contact with each outcome, using
effect sizes from all studies that measured both predictors (Figure 8a).
By using partial correlations, we estimated the effect of one form of
contact while controlling for the other. Across 105,374 participants
from 27 samples in 16 studies, we found stronger evidence for a positive
association between negative contact and perceived injustice
(\(r = .14, [.05, .23]\); \(\Pr (r > 0) = 99.8\%\)) than for a negative
association between positive contact and perceived injustice
(\(r = -.05, [-.11, .01]\); \(\Pr (r < 0) = 94.8\%\)). Across 107,451
participants from 16 samples in 9 studies, we again found stronger
evidence for a positive association between negative contact and
collective action (\(r = .11, [.01, .20]\); \(\Pr (r > 0) = 98.4\%\))
than for a negative association between positive contact and collective
action (\(r = .00, [-.03, .06]\); \(\Pr (r < 0) = 40.7\%\)). Across
8,273 participants from 7 samples in 3 studies, we also found stronger
evidence for a positive association between negative contact and policy
support (\(r = .04, [.00, .07]\); \(\Pr (r > 0) = 98.3\%\)) than for a
negative association between positive contact and policy support
(\(r = .03, [-.03, .10]\); \(\Pr (r < 0) = 12.7\%\)). Across 110,257
participants from 29 samples in 17 studies, we found a negative
association between positive and negative contact
(\(r = -.16, [-.25, -.05]\); \(\Pr (r < 0) = 99.8\%\)).

@@

A combination of moderators explained about a third of the between-study
variance in the association between intergroup contact and perceived
injustice. We found that, on average, this association was negative only
in studies of adults that measured contact \emph{directly} (e.g., by
asking about the number of friends from the advantaged outgroup). This
suggests that, as hypothesized, the `ironic' effects result from direct
contact and not from other contextual factors. We also found that, on
average, effect sizes were greater for studies in post-colonial settings
or on short-term migration than for studies in other contexts. Future
research should systematically investigate variance in how contact
affects support for social change across settings and cultures.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-becker_friend_2013}{}%
Becker, J. C., Wright, S. C., Lubensky, M. E., \& Zhou, S. (2013).
Friend or ally: Whether cross-group contact undermines collective action
depends on what advantaged group members say (or don't say).
\emph{Personality and Social Psychology Bulletin}, \emph{39}(4),
442--455. \url{https://doi.org/10.1177/0146167213477155}

@@

@@

@@

\leavevmode\hypertarget{ref-cakal_investigation_2011}{}%
Çakal, H., Hewstone, M., Schwär, G., \& Heath, A. (2011). An
investigation of the social identity model of collective action and the
``sedative'' effect of intergroup contact among Black and White students
in South Africa. \emph{British Journal of Social Psychology},
\emph{50}(4), 606--627.
\url{https://doi.org/10.1111/j.2044-8309.2011.02075.x}

@@

\leavevmode\hypertarget{ref-dixon_intergroup_2007}{}%
Dixon, J., Durrheim, K., \& Tredoux, C. (2007). Intergroup contact and
attitudes toward the principle and practice of racial equality.
\emph{Psychological Science}, \emph{18}(10), 867--872.
\url{https://doi.org/10.1111/j.1467-9280.2007.01993.x}

\leavevmode\hypertarget{ref-dixon_paradox_2010}{}%
Dixon, J., Durrheim, K., Tredoux, C., Tropp, L., Clack, B., \& Eaton, L.
(2010). A paradox of integration? Interracial contact, prejudice
reduction, and perceptions of racial discrimination: A paradox of
integration. \emph{Journal of Social Issues}, \emph{66}(2), 401--416.
\url{https://doi.org/10.1111/j.1540-4560.2010.01652.x}

@@

\leavevmode\hypertarget{ref-droogendyk_renewed_2016}{}%
Droogendyk, L., Louis, W. R., \& Wright, S. C. (2016). Renewed promise
for positive cross-group contact: The role of supportive contact in
empowering collective action. \emph{Canadian Journal of Behavioural
Science / Revue Canadienne Des Sciences Du Comportement}, \emph{48}(4),
317--327. \url{https://doi.org/10.1037/cbs0000058}

@@

@@

\leavevmode\hypertarget{ref-hayward_how_2018}{}%
Hayward, L. E., Tropp, L. R., Hornsey, M. J., \& Barlow, F. K. (2018).
How negative contact and positive contact with Whites predict collective
action among racial and ethnic minorities. \emph{British Journal of
Social Psychology}, \emph{57}(1), 1--20.
\url{https://doi.org/10.1111/bjso.12220}

\leavevmode\hypertarget{ref-hassler_large-scale_2020}{}%
Hässler, T., Ullrich, J., Bernardino, M., Shnabel, N., Laar, C. V.,
Valdenegro, D., Sebben, S., Tropp, L. R., Visintin, E. P., González, R.,
Ditlmann, R. K., Abrams, D., Selvanathan, H. P., Branković, M., Wright,
S., Zimmermann, J. von, Pasek, M., Aydin, A. L., Žeželj, I., \ldots{}
Ugarte, L. M. (2020). A large-scale test of the link between intergroup
contact and support for social change. \emph{Nature Human Behaviour},
\emph{4}(4), 380--386. \url{https://doi.org/10.1038/s41562-019-0815-z}

@@

\leavevmode\hypertarget{ref-koschate_when_2012}{}%
Koschate, M., Hofmann, W., \& Schmitt, M. (2012). When East meets West:
A longitudinal examination of the relationship between group relative
deprivation and intergroup contact in reunified Germany. \emph{British
Journal of Social Psychology}, \emph{51}(2), 290--311.
\url{https://doi.org/10.1111/j.2044-8309.2011.02056.x}

>>>>

\noindent Growing evidence suggests that intergroup contact, psychology's
most-researched paradigm for reducing prejudice, has the `ironic' effect
of reducing support for social change in disadvantaged groups. We
conducted a preregistered meta-analytic test of this effect across 98
studies with 140 samples of 213,085 disadvantaged-group members. As
predicted, intergroup contact was, on average, associated with less
perceived injustice (\(r = -.07\)), collective action (\(r = -.06\)),
and support for reparative policies (\(r = -.07\)). However, these
associations were small, variable, and consistent with alternative
explanations. Moderator analyses found that associations were strongest
in studies of adults that measured intergroup contact directly and
examined short-term migration or post-colonial intergroup relations.
After controlling for the positive association of negative contact with
support for social change, positive contact was not associated with any
outcome in studies that measured both predictors. We close by
highlighting open questions about the relationship between intergroup
contact and social change.

@@

It is ironic that psychology's most-researched paradigm for reducing
prejudice---intergroup contact---should hinder, rather than help, social
change. Hundreds of studies have confirmed that intergroup contact is
associated with less negative feelings toward outgroup members (for a
meta-analysis, see Pettigrew \& Tropp, 2006). Most of these studies have
focused on advantaged-group members' contact experiences with
disadvantaged-group members (Tropp \& Pettigrew, 2005), guided by the
idea that reducing advantaged-group members' prejudice will prevent
interpersonal discrimination. Reducing prejudice, however, is rarely
enough to overcome institutional and structural discrimination. Instead,
as the 2020 Black Lives Matter protests have shown, social change often
requires political mobilization by the disadvantaged. Critics have
argued that, by fostering harmony and reducing conflict, intergroup
contact might distract disadvantaged-group members from the injustice
they face and, thereby, undermine their motivation for social change
(Dixon et al., 2005; Reicher, 2007; Wright, 2003). By reducing
disadvantaged-group members' opposition to injustice, intergroup contact
might thus have the `ironic' (Saguy et al., 2009), `sedative' (Çakal et
al., 2011), or `paradoxical' (Dixon et al., 2010) effect of hindering
social change.

@@

Collective action is any action directed at improving the conditions of
the disadvantaged ingroup (Wright et al., 1990). Collective action is
crucial to achieving social change because advantaged-group members
rarely give up their advantages without sustained pressure from
disadvantaged-group members (Blumer, 1958; for a review, see Dixon et
al., 2012). Researchers have argued that intergroup contact could
discourage disadvantaged-group members from engaging in collective
action by diminishing their perceived discrimination (Tropp et al.,
2012), by blurring boundaries between the disadvantaged ingroup and the
advantaged outgroup (Saguy et al., 2009), by reducing negative attitudes
toward the advantaged outgroup (Wright \& Lubensky, 2009), and by
quelling their anger (Hayward et al., 2018; Tausch et al., 2015). To the
extent that it discourages collective action, intergroup contact reduces
support for social change.

@@

Figure 1 shows a flow diagram illustrating our search strategy, study
selection, and data collection. Our preregistered search strategy
returned 2,610 unique records from electronic databases. Of these, we
excluded 2,379 (91\%) ineligible records after screening titles,
abstracts, and keywords. Of the remaining 231 records, we excluded 116
(50\%) ineligible records after reviewing full-text manuscripts. We
supplemented these records with 15 unpublished studies and 6 records
that cited at least three relevant works. Of 136 eligible studies, we
had to exclude 38 (28\%) studies for which we could not extract or
impute any relevant effect size. At each stage, we also excluded studies
that used the same data as another study. Our final sample comprised
effect sizes from \(J = 98\) studies spanning \(N = 213,085\)
participants in \(I = 140\) samples (for a complete list of studies, see
SOM-U).

Figure 2 provides a qualitative overview of the literature. Even though
most participants came from studies conducted in India (\(N = 104,639\))
and the United States (\(N = 60,983\); Figure 2a), most samples were
collected in North America (\(I = 48\)) and Europe (\(I = 43\)), with
few samples from South America and Africa (Figure 2b). Most samples
focused on relative inequalities that resulted from long-term migration
(\(I = 59\)), slavery (\(I = 19\)), colonization (\(I = 19\)), and
short-term migration (\(I = 14\)). With few exceptions, studies used
observational, cross-sectional survey designs. Samples sizes ranged from
64 to 49,764 (\(\textit{Mdn} = 287\)). Unlike most psychological
research, 2 in 5 samples used probability or representative sampling. Of
all samples, 55\% (\(I = 77\)) were collected with the intention to
examine the effects of intergroup contact on the outcomes considered in
this meta-analysis. Many studies instead examined acculturation
processes (Berry, 1997) and measured intergroup contact and perceived
injustice without hypothesizing that the former would influence the
latter. Together, these observations highlight both limitations and
strengths of the empirical literature.

@@

\textbf{\emph{Perceived Injustice.}} Across 203,637 participants from
126 samples in 86 studies, we found strong evidence for a weak
association (\(r = -.07, [-.10, -.04]\)) between intergroup contact and
perceived injustice, with \(>99.9\%\) of posterior samples for the mean
correlation coefficient falling below zero. We found evidence that
correlation coefficients varied across studies
(\(\tau_J = .14, [.11, .17]\)) and across samples within studies
(\(\tau_I = .08, [.05, .12]\)). Based on these analyses, we predicted
that 80\% of studies would result in correlation coefficients between
-.25 and .11 and that researchers would need sample sizes of at least
2,542 participants to find significant associations (\(\alpha = .05\),
two-sided) in 80\% of their studies.\footnote{Sample sizes are based on
  posterior predictions which implied that, for 80\% of studies, the
  absolute correlation coefficient would be \(|r| > .039\) for perceived
  injustice, \(|r| > .045\) for collective action, and \(|r| > .035\)
  for policy support.}

@@

\begin{figure*}
\centering
\caption{Estimated effect size as a function of cultural distance from the United States, with point estimates and uncertainty intervals for each country}
\includegraphics[scale=1]{../figures/figure-5}
\caption*{\textit{Note.} We had data from too few countries to reach firm conclusions about the direction of these associations for perceived injustice ($\beta = -0.38, [-1.45, 0.72]$; $\Pr (\beta < 0) = 75.7\%$), collective action ($\beta = -0.09, [-1.60, 1.37]$; $\Pr (\beta < 0) = 55.1\%$), and policy support ($\beta = -0.92, [-2.13, 0.55]$; $\Pr (\beta < 0) = 90.3\%$).}
\label{fig:f5}
\end{figure*}
\begin{figure*}
\centering
\caption{Results from the random-effects meta-regression tree analysis}
\includegraphics[scale=1]{../figures/figure-6}
\caption*{\textit{Note.} Posterior distributions for the estimated correlation coefficient in each leaf of the meta-regression tree, highlighting the proportion of posterior samples for which $r_\text{mean} < 0$.  S.-T. Migration = Short-Term Migration.}
\label{fig:f6}
\end{figure*}

@@

Second, we used meta-regression trees to discover interactions between
moderator variables that best explained heterogeneity in effect sizes
(Li et al., 2020). Figure 6 shows the resulting meta-regression that
explained more variance across samples than any individual moderator
(\(R^2 = 31\%\)). We found that intergroup contact was associated with
less perceived injustice only in studies that focused on adults and that
measured intergroup contact directly. Among these studies, this
association was stronger in settings in which the groups' relative
inequality stemmed from short-term migration or colonization
(\(r = -.18, [-.23, -.13]\)) than in other settings
(\(r = -.09, [-.13, -.06]\)).

@@

We ran three random-effects meta-analysis models estimating the partial
correlations of positive and negative contact with each outcome, using
effect sizes from all studies that measured both predictors (Figure 8a).
By using partial correlations, we estimated the effect of one form of
contact while controlling for the other. Across 106,270 participants
from 29 samples in 17 studies, we found somewhat stronger evidence for a
positive association between negative contact and perceived injustice
(\(r = .14, [.06, .23]\); \(\Pr (r > 0) = 99.9\%\)) than for a negative
association between positive contact and perceived injustice
(\(r = -.05, [-.10, .01]\); \(\Pr (r < 0) = 95.6\%\)). Across 107,451
participants from 16 samples in 9 studies, we again found stronger
evidence for a positive association between negative contact and
collective action (\(r = .11, [.01, .20]\); \(\Pr (r > 0) = 98.4\%\))
than for a negative association between positive contact and collective
action (\(r = .00, [-.03, .06]\); \(\Pr (r < 0) = 40.7\%\)). Across
8,273 participants from 7 samples in 3 studies, we also found stronger
evidence for a positive association between negative contact and policy
support (\(r = .04, [.00, .07]\); \(\Pr (r > 0) = 98.3\%\)) than for a
negative association between positive contact and policy support
(\(r = .03, [-.03, .10]\); \(\Pr (r < 0) = 12.7\%\)). Across 111,153
participants from 31 samples in 18 studies, we found a negative
association between positive and negative contact
(\(r = -.16, [-.25, -.07]\); \(\Pr (r < 0) = 99.8\%\)).

@@

A combination of moderators explained about a third of the between-study
variance in the association between intergroup contact and perceived
injustice. We found that, on average, this association was negative only
in studies of adults that measured contact \emph{directly} (e.g., by
asking about the number of friends from the advantaged outgroup). This
suggests that, as hypothesized, the `ironic' effects result from direct
contact and not from other contextual factors. We also found that, on
average, effect sizes were greater for studies in (post-)colonial
settings or on short-term migration than for studies in other contexts.
Future research should systematically investigate variance in how
contact affects support for social change across settings and cultures.

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-3201}{}%
*Albzour, M., Penic, S., Nasser, R., \& Green, E. G. T. (2019). Support
for ``normalization'' of relations between Palestinians and Israelis,
and how it relates to contact and resistance in the West Bank.
\emph{Journal of Social and Political Psychology}, \emph{7}(2),
978--996. \url{https://doi.org/10.5964/jspp.v7i2.877}

\leavevmode\hypertarget{ref-2023}{}%
*Bagci, S. C. (2014). \emph{Cross-ethnic friendships in multiethnic
educational settings: Consequences for psychological, academic and
motivational outcomes among young adolescents} {[}Doctoral dissertation,
Goldsmiths, University of London{]}.
\url{https://doi.org/10.25602/GOLD.00011030}

\leavevmode\hypertarget{ref-2395}{}%
*Bagci, S. C., \& Stathi, S. (2020). \emph{Secure and insecure ingroup
identification as mediators between identity threat and psychological
well-being and intergroup bias} {[}Manuscript in preparation{]}. Faculty
of Arts and Social Sciences, Sabanci University.

\leavevmode\hypertarget{ref-423}{}%
*Bagci, S. C., Stathi, S., \& Piyale, Z. E. (2019). Imagined contact
facilitates acculturation, sometimes: Contradicting evidence from two
sociocultural contexts. \emph{Cultural Diversity and Ethnic Minority
Psychology}, \emph{25}(4), 539--552.
\url{https://doi.org/10.1037/cdp0000256}

\leavevmode\hypertarget{ref-322}{}%
*Bagci, S. C., \& Turnuklu, A. (2019). Intended, unintended, and unknown
consequences of contact: The role of positive-negative contact on
outgroup attitudes, collective action tendencies, and psychological
well-being. \emph{Social Psychology}, \emph{50}(1), 7--23.
\url{https://doi.org/10.1027/1864-9335/a000355}

\leavevmode\hypertarget{ref-414}{}%
*Bagci, S. C., Turnuklu, A., \& Bekmezci, E. (2018a). Cross-group
friendships and psychological well-being: A dual pathway through social
integration and empowerment. \emph{British Journal of Social
Psychology}, \emph{57}(4), 773--792.
\url{https://doi.org/10.1111/bjso.12267}

\leavevmode\hypertarget{ref-2257}{}%
*Bagci, S. C., Turnuklu, A., \& Bekmezci, E. (2018b). The buffering role
of in-group identification and intergroup contact on the association
between perceived discrimination and mental health. \emph{Journal of
Community \& Applied Social Psychology}, \emph{28}(5), 293--305.
\url{https://doi.org/10.1002/casp.2357}

\leavevmode\hypertarget{ref-2396}{}%
*Bagci, S. C., Verkuyten, M., Koc, Y., Turnuklu, A., Piyale, Z. E., \&
Bekmezci, E. (2020). Being tolerated and being discriminated against:
Links to psychological well-being through threatened social identity
needs. \emph{European Journal of Social Psychology}, \emph{50}(7),
1463--1477. \url{https://doi.org/10.1002/ejsp.2699}

\leavevmode\hypertarget{ref-2075}{}%
*Becker, J. C., Wright, S. C., Lubensky, M. E., \& Zhou, S. (2013).
Friend or ally: Whether cross-group contact undermines collective action
depends on what advantaged group members say (or don't say).
\emph{Personality and Social Psychology Bulletin}, \emph{39}(4),
442--455. \url{https://doi.org/10.1177/0146167213477155}

\leavevmode\hypertarget{ref-1045}{}%
*Benner, A. D., \& Wang, Y. (2017). Racial/ethnic discrimination and
adolescents' well-being: The role of cross-ethnic friendships and
friends' experiences of discrimination. \emph{Child Development},
\emph{88}(2), 493--504. \url{https://doi.org/10.1111/cdev.12606}

@@

\leavevmode\hypertarget{ref-2397}{}%
*Bond, J., \& Dixon, J. (2020). \emph{White and Black UK residents'
support for collective action for the removal of public symbols
associated with historical racism in England} {[}Unpublished dataset{]}.
The Open University.

@@

\leavevmode\hypertarget{ref-2398}{}%
*Bracegirdle, C. (2020). \emph{Social network study of two diverse
schools in Northern England} {[}Unpublished dataset{]}. University of
Oxford.

@@

\leavevmode\hypertarget{ref-480}{}%
*Castro, V. S. (2005). Discriminación percibida y autoestima en jóvenes
de grupos minoritarios y mayoritarios en Costa Rica {[}Perceived
discrimination and self-esteem among ethnic majority and minority youths
in Costa Rica{]}. \emph{Revista Interamericana de Psicología},
\emph{39}(1), 93--106.

\leavevmode\hypertarget{ref-1914}{}%
*Castro Solano, A., \& Lupano Perugini, M. L. (2013). Predictores de la
adaptación sociocultural de estudiantes universitarios extranjeros en
Argentina {[}Predictors of socio-cultural adaptation of foreign
university students in Argentina{]}. \emph{Interdisciplinaria},
\emph{30}(2), 265--281.

\leavevmode\hypertarget{ref-3205}{}%
*Cernat, V. (2019). Interminority contact and solidarity under conflict:
Evidence from the Hungarian and Roma minorities in Romania. \emph{Basic
and Applied Social Psychology}, \emph{41}(5), 326--339.
\url{https://doi.org/10.1080/01973533.2019.1652179}

\leavevmode\hypertarget{ref-91}{}%
*Chang, T. K. (2017). \emph{Psychological distress of Asian
international students: Applying intergroup contact theory} {[}Doctoral
dissertation, Indiana University{]}. ProQuest Dissertations and Theses
Global.

\leavevmode\hypertarget{ref-1163}{}%
*Cruse, S. L. (2015). \emph{The relationship of ethnic identity,
perceived discrimination, and outgroup contact to racial attitudes
toward White people in African American college students} {[}Doctoral
dissertation, University of Kansas{]}. ProQuest Dissertations and Theses
Global.

\leavevmode\hypertarget{ref-1248}{}%
*Çakal, H., Hewstone, M., Güler, M., \& Heath, A. (2016). Predicting
support for collective action in the conflict between Turks and Kurds:
Perceived threats as a mediator of intergroup contact and social
identity. \emph{Group Processes \& Intergroup Relations}, \emph{19}(6),
732--752. \url{https://doi.org/10.1177/1368430216641303}

\leavevmode\hypertarget{ref-284}{}%
*Çakal, H., Hewstone, M., Schwär, G., \& Heath, A. (2011). An
investigation of the social identity model of collective action and the
``sedative'' effect of intergroup contact among Black and White students
in South Africa. \emph{British Journal of Social Psychology},
\emph{50}(4), 606--627.
\url{https://doi.org/10.1111/j.2044-8309.2011.02075.x}

\leavevmode\hypertarget{ref-1614}{}%
*Çelebi, E., Verkuyten, M., \& Smyrnioti, N. (2016). Support for Kurdish
language rights in Turkey: The roles of ethnic group, group
identifications, contact, and intergroup perceptions. \emph{Ethnic and
Racial Studies}, \emph{39}(6), 1034--1051.
\url{https://doi.org/10.1080/01419870.2015.1103881}

\leavevmode\hypertarget{ref-783}{}%
*de Vroome, T., Verkuyten, M., \& Martinovic, B. (2014). Host national
identification of immigrants in the Netherlands. \emph{International
Migration Review}, \emph{48}(1), 1--27.
\url{https://doi.org/10.1111/imre.12063}

\leavevmode\hypertarget{ref-3196}{}%
*Di Bernardo, G. A., Vezzali, L., Stathi, S., McKeown Jones, S., Cocco,
V. M., Saguy, T., \& Dixon, J. (2019). Fostering social change among
advantaged and disadvantaged group members: Integrating intergroup
contact and social identity perspectives on collective action.
\emph{Group Processes \& Intergroup Relations}.
\url{https://doi.org/10.1177/1368430219889134}

\leavevmode\hypertarget{ref-1807}{}%
*Di Saint Pierre, F., Martinovic, B., \& de Vroome, T. (2015). Return
wishes of refugees in the Netherlands: The role of integration, host
national identification and perceived discrimination. \emph{Journal of
Ethnic and Migration Studies}, \emph{41}(11), 1836--1857.
\url{https://doi.org/10.1080/1369183X.2015.1023184}

\leavevmode\hypertarget{ref-93}{}%
*Dixon, J., Çakal, H., Khan, W., Osmany, M., Majumdar, S., \& Hassan, M.
(2017). Contact, political solidarity and collective action: An Indian
case study of relations between historically disadvantaged communities.
\emph{Journal of Community \& Applied Social Psychology}, \emph{27}(1),
83--95. \url{https://doi.org/10.1002/casp.2296}

@@

\leavevmode\hypertarget{ref-1386}{}%
*Dixon, J., Durrheim, K., \& Tredoux, C. (2007). Intergroup contact and
attitudes toward the principle and practice of racial equality.
\emph{Psychological Science}, \emph{18}(10), 867--872.
\url{https://doi.org/10.1111/j.1467-9280.2007.01993.x}

\leavevmode\hypertarget{ref-336}{}%
*Dixon, J., Durrheim, K., Tredoux, C., Tropp, L. R., Clack, B., \&
Eaton, L. (2010). A paradox of integration? Interracial contact,
prejudice reduction, and perceptions of racial discrimination.
\emph{Journal of Social Issues}, \emph{66}(2), 401--416.
\url{https://doi.org/10.1111/j.1540-4560.2010.01652.x}

@@

\leavevmode\hypertarget{ref-2033}{}%
*Droogendyk, L., Louis, W. R., \& Wright, S. C. (2016). Renewed promise
for positive cross-group contact: The role of supportive contact in
empowering collective action. \emph{Canadian Journal of Behavioural
Science / Revue Canadienne Des Sciences Du Comportement}, \emph{48}(4),
317--327. \url{https://doi.org/10.1037/cbs0000058}

\leavevmode\hypertarget{ref-4001}{}%
*Eller, A., Abrams, D., Wright, S. C., \& Davies, B. (2020). Effects of
intergroup contact and relative gratification vs deprivation on
prejudice on both sides of the U.S./Mexico status divide. \emph{Journal
of Applied Social Psychology}, \emph{50}(7), 406--417.
\url{https://doi.org/10.1111/jasp.12669}

\leavevmode\hypertarget{ref-1549}{}%
*Forbes, J. M. (2001). \emph{Affirmative action attitudes in black and
white} {[}Doctoral dissertation, Wayne State University{]}. ProQuest
Dissertations and Theses Global.

\leavevmode\hypertarget{ref-325}{}%
*Franco, M. (2019). Let the racism tell you who your friends are: The
effects of racism on social connections and life-satisfaction for
Multiracial people. \emph{International Journal of Intercultural
Relations}, \emph{69}, 54--65.
\url{https://doi.org/10.1016/j.ijintrel.2018.12.005}

@@

\leavevmode\hypertarget{ref-1576}{}%
*Goto, S. G., Gee, G. C., \& Takeuchi, D. T. (2002). Strangers still?
The experience of discrimination among Chinese Americans. \emph{Journal
of Community Psychology}, \emph{30}(2), 211--224.
\url{https://doi.org/10.1002/jcop.9998}

\leavevmode\hypertarget{ref-1949}{}%
*Haase, A., Rohmann, A., \& Hallmann, K. (2019). An ecological approach
to psychological adjustment: A field survey among refugees in Germany.
\emph{International Journal of Intercultural Relations}, \emph{68},
44--54. \url{https://doi.org/10.1016/j.ijintrel.2018.10.003}

@@

\leavevmode\hypertarget{ref-4003}{}%
*Hayward, L. E., Hornsey, M. J., Tropp, L. R., \& Barlow, F. K. (2017).
Positive and negative intergroup contact predict Black and White
Americans' judgments about police violence against Black Americans.
\emph{Journal of Applied Social Psychology}, \emph{47}(11), 605--615.
\url{https://doi.org/10.1111/jasp.12463}

\leavevmode\hypertarget{ref-2309}{}%
*Hayward, L. E., Tropp, L. R., Hornsey, M. J., \& Barlow, F. K. (2018).
How negative contact and positive contact with Whites predict collective
action among racial and ethnic minorities. \emph{British Journal of
Social Psychology}, \emph{57}(1), 1--20.
\url{https://doi.org/10.1111/bjso.12220}

\leavevmode\hypertarget{ref-3054}{}%
*Hässler, T., Ullrich, J., Bernardino, M., Shnabel, N., Laar, C. V.,
Valdenegro, D., Sebben, S., Tropp, L. R., Visintin, E. P., González, R.,
Ditlmann, R. K., Abrams, D., Selvanathan, H. P., Branković, M., Wright,
S., Zimmermann, J. von, Pasek, M., Aydin, A. L., Žeželj, I., \ldots{}
Ugarte, L. M. (2020). A large-scale test of the link between intergroup
contact and support for social change. \emph{Nature Human Behaviour},
\emph{4}(4), 380--386. \url{https://doi.org/10.1038/s41562-019-0815-z}

\leavevmode\hypertarget{ref-1407}{}%
*Head, R. N. (2013). \emph{Coping with perceived discrimination}
{[}Doctoral dissertation, North Carolina State University{]}. ProQuest
Dissertations and Theses Global.

@@

\leavevmode\hypertarget{ref-1494}{}%
*Herda, D. E. (2013). \emph{Hostile contact: Self-reports of
interpersonal discrimination among Blacks, Whites and Latinos in
Chicago} {[}Doctoral dissertation, University of California, Davis{]}.
ProQuest Dissertations and Theses Global.

\leavevmode\hypertarget{ref-324}{}%
*Herda, D. E. (2016). The specter of discrimination: Fear of
interpersonal racial discrimination among adolescents in Chicago.
\emph{Social Science Research}, \emph{55}, 48--62.
\url{https://doi.org/10.1016/j.ssresearch.2015.09.010}
